---
title: "Read"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: FALSE
    code_folding: show
    number_sections: TRUE
---

# SET-UP

```{r markdown, echo=F}
library(knitr)
opts_chunk$set(echo = T, collapse = T, fig.align = "center", fig.width = 9, fig.height = 6)
options(width = 200)
```    

```{r settings, message = F}
source('0.0_settings.R')
```

# Landings
## ZIFF
```{r ziffsp, message = F}
sp_code <- find.species(my.species,'ziff')
kable(sp_code) 
```

```{r ziff, message = F}
ziffR <- paste0(dir.rdat, "ziff.Rdata")

if(!file.exists(ziffR)){
    ziff <- read.ziff(sp_code,path=dir.ziff,year=my.year,language='en')
    
    # Species included:
    with(ziff, table(species,year))
    ziff <- ziff[ziff$species==250,]   # code 960: few heads in 1996  code 951: wrong whelk

    save(ziff, file = ziffR)
}else{
    load(ziffR)
}

```
### CHECKS {.tabset}

#### areas
Nafo 5Y and 5Ze are partially in the US.
```{r ziff areas}
# NAFO 5 = landings in Canadian EEZ (northern contingent)
kable(with(ziff, table(nafo,month)))     
```

#### NA weights
How many NA weights are there?
```{r ziff nas}
na <- cbind(all=with(ziff,table(year,useNA = 'always')),
                na=with(ziff[is.na(ziff$catch),],table(year,useNA = 'always')))
kable(na)
```

#### 0 weights
How many 0 weights are there?
```{r ziff zeros}
with(ziff[!is.na(ziff$catch) & ziff$catch==0,],table(year))
```

#### Jan-1
Check that there is no abnormal quantity of landings on January first (possibly records for which data is unknown).
```{r ZIFF_jan1}
ggplot(ziff,aes(x=lubridate::yday(capt.date)))+geom_histogram(binwidth = 1)+
    scale_x_continuous(expand=c(0,0))+
    scale_y_continuous(expand=c(0,0))
```

#### NPBC
There are 3 special nbpc numbers that indicate landings that are not necessarily daily (e.g., monthly totals) or for which the data is only approximate. For instance, B00000 landings seems to be on the first of last of a certain month. For analyses of daily landings these might need to be excluded.
```{r ZIFF_npbc}
kable(with(ziff[ziff$npbc %in% c('000000', 'B00000', '190000'),], table(year,npbc)))  
```

## NAFO

```{r nafo, message = F}
nafoAR <- paste0(dir.rdat, "nafoA.Rdata")
nafoBR <- paste0(dir.rdat, "nafoB.Rdata")

if(!all(file.exists(c(nafoAR,nafoBR)))){
    ### NAFO database A (annual) ---------------------------------
    sp_code <- find.species(my.species,'nafoA');sp_code
    nafoA <- read.nafoA(paste0(dir.nafo,'statlandA.csv'),species=sp_code[1], year=my.year) 

    # Canadian caught mackerel outside of my.nafo
    with(nafoA[!grepl(paste0(my.nafo, collapse = "|"),nafoA$nafo)&
                 grepl('canada',nafoA$country,ignore.case = TRUE),] ,table(year,nafo))

    # only keep my.nafo; other areas are NOT from nothern contingent (winter months in US or Europe)
    nafoA <- nafoA[grepl(paste0(my.nafo, collapse = "|"),nafoA$nafo),] 
    table(nafoA$nafo)
    
    save(nafoA, file = nafoAR)
    
    ### NAFO data base B (monthly) ---------------------------------
    sp_code <- find.species(my.species,'nafoB');sp_code
    nafoB <- read.nafoB(path=dir.nafo,species=sp_code[1],year=my.year,overwrite=FALSE)  # overwrite=TRUE if new download   

    # Canadian caught mackerel outside of my.nafo
    with(nafoB[!grepl(paste0(my.nafo, collapse = "|"),nafoB$nafo)&
                 grepl('canada',nafoB$country,ignore.case = TRUE),] ,table(year,nafo))
    with(nafoB[!grepl(paste0(my.nafo, collapse = "|"),nafoB$nafo)&
                 grepl('canada',nafoB$country,ignore.case = TRUE),] ,table(month,nafo))
    
    # only keep my.nafo  (though 5Ze in summer; wouldn't they be northern contingent fish??)
    nafoB <- nafoB[grepl(paste0(my.nafo, collapse = "|"),nafoB$nafo),] 
    table(nafoB$nafo)
    
    # remove useless columns 
    nafoB <- nafoB[,c('year','month','country','nafo','gear','gear.cat','tonnage.class','prov','catch')]

    save(nafoB, file = nafoBR)
}else{
    load(nafoAR)
    load(nafoBR)
}
```
### CHECK {.tabset}
Nafo B is used for landings prior to 1995.

#### compare
```{r NAFO_comparison}
ggplot()+
    geom_bar(data=nafoA,aes(x=year,y=catch,fill="Nafo A"),stat='identity',alpha=0.4)+
    geom_bar(data=nafoB,aes(x=year,y=catch,fill="Nafo B"),stat='identity',alpha=0.4)+
    geom_vline(xintercept=1985)+
    labs(fill='',y='Landings (t)',x='Year')+
    scale_x_continuous(expand=c(0,0))+ scale_y_continuous(expand=c(0,0))
```

## CATCH = NAFO + ZIFF

```{r CATCH, message = F}
## bind both
ziff$source <-'ziff'
ziff$catch <- ziff$catch/1000            # kg to tonnes
ziff$country <- 'Canada'                 # consistency with nafo
ziff$prov <- ifelse(ziff$prov.land=='Unknown',as.character(ziff$prov.home),as.character(ziff$prov.land))

nafoB$source <- 'nafo'                   # work with detailed B base

sel <- c('year','month','nafo','gear.cat','source','catch','country','prov')
catch <- rbind(ziff[ziff$year>=1995,sel],        # ziff from 1995 onwards
               nafoB[nafoB$year<1995,sel],       # nafoB before
               nafoB[nafoB$year>=1995 & nafoB$country!='Canada',sel]) # foreign post 1994

save(catch, file = paste0(dir.rdat, "catch.Rdata"))

plotCatch(catch,x='year',y='catch')
```
